The \dealii{} Library, Version 9.2

### 2020-06-01 9.2/paper.tex
Minor edits.

As part of this effort, we ran benchmarks on the TACC Frontera system, where we were able to apply the matrix-free geometric multigrid framework to a variable viscosity Stokes system and achieved weak and strong scaling up to 114k MPI ranks with up to
\num{2.1e11} unknowns. This is likely the currently largest block system solved with \dealii{} and required various optimizations
and fixes on top of the ones mentioned above:
1) Bug fixes to concurrent point to point communications.
2) Fixes multigrid transfer with adaptive refinement and more than \num{4e9} unknowns.
3) Fixes to index sets in block indices with more than \num{4e9} unknowns.
4) Fixes to computations with more than \num{4e9} active cells.
5) Implementation of IDR(s) solvers to reduce memory overhead.
For more details, see \cite{clevenger_stokes19}.

@@

In the previous release we added support for symbolic expressions, leveraging the
SymEngine library \cite{symengine-web-page}.
Although effective, evaluating lengthy expressions could be a bottle-neck as this
was performed using dictionary-based substitution.
We have improved on this by implementing a \texttt{BatchOptimizer} class in the 
namespace \texttt{Differentiation::SD} that collects several \texttt{Expression}s
and transforms them in such a way that the equivalent result is returned through
a quicker code path.
This may be done by simply using common subexpression elimination (CSE) for the
dictionary-based expressions, by transformation to a set of nested 
\texttt{std::function}s (the equivalent to \texttt{SymPy}'s "lambdify", with or
without using CSE), or by offloading to these expressions to the \texttt{LLVM}
just-in-time (JIT) compiler.
Although, each of these features is implemented and tested in the SymEngine
library itself, the \texttt{BatchOptimizer} class provides both uniform
interface to their classes and a convenient interface for scalar expressions,
as well as tensorial expressions formed using the \texttt{deal.II} tensor and 
symmetric tensor classes.
It, like the \texttt{Expression} class, is also serializable.

The way that the batch optimizer may be employed within a user's code is shown
in the pseudo-code below.
As per usual, one would first define some independent variables, and 
subsequently compute some symbolic expressions that are dependent on these 
independent variables. 
These expression could be, for example, scalar  expressions or tensors of 
expressions.
Instead of evaluating these expressions directly, the user would now create an
optimizer to evaluate the dependent functions.
In this example, the selected arithmetic type numerical result will be of type
\texttt{double}, and the \texttt{LLVM} JIT optimizer will be invoked.
It will employ common subexpression elimination and aggressive optimizations
during compilation.
The user then informs the optimizer of all of the independent variables and the
dependent expressions, and invokes the optimization process.
This is an expensive call, as it determines an equivalent code path to evaluate
all of the dependent functions at once.
However, in many cases each evaluation has significantly less computational cost
than evaluating the symbolic expressions directly.
Evaluation is performed when the user constructs a substitution map, giving each
independent variable a numerical representation, and passes those to the 
optimizer.
After this step, the numerical equivalent of the individual dependent expressions
may finally be retreived from the optimizer.

@@

  \item \texttt{step-50} is a program that demonstrates the parallel geometric multigrid features
  in \dealii{} as described in \cite{ClevengerHeisterKanschatKronbichler2019}.
  The problem considered is a variable viscosity Laplace equation and it is solved
  with three different approaches:
  1. With a matrix-based geometric multigrid based on Trilinos or PETSc. 2. With a matrix-free
  geometric multigrid. 3. Using algebraic multigrid (Trilinos ML).

>>>>

As part of this effort, we ran benchmarks on the TACC Frontera system, where we were able to apply the matrix-free geometric multigrid framework to a variable viscosity Stokes system and achieved weak and strong scaling up to 114k MPI ranks with up to
\num{2.1e11} unknowns. This is likely the currently largest block system solved with \dealii{} and required various optimizations
and fixes on top of the ones mentioned above:
(i) Bug fixes to concurrent point to point communications.
(ii) Fixes multigrid transfer with adaptive refinement and more than \num{4e9} unknowns.
(iii) Fixes to index sets in block indices with more than \num{4e9} unknowns.
(iv) Fixes to computations with more than \num{4e9} active cells.
(v) Implementation of IDR(s) solvers to reduce memory overhead.
For more details, see \cite{clevenger_stokes19}.

@@

In the previous release we added support for symbolic expressions, leveraging the
SymEngine library \cite{symengine-web-page}.
Although effective, evaluating lengthy expressions could be a bottleneck as this
was performed using dictionary-based substitution.
We have improved on this by implementing a \texttt{BatchOptimizer} class in the 
namespace \texttt{Differentiation::SD} that collects several \texttt{Expression}s
and transforms them in such a way that the equivalent result is returned through
a quicker code path.
This may be done by simply using common subexpression elimination (CSE) for the
dictionary-based expressions, by transformation to a set of nested 
\texttt{std::function} objects (the equivalent to \texttt{SymPy}'s ``lambdify'', with or
without using CSE), or by offloading these expressions to the \texttt{LLVM}
just-in-time (JIT) compiler.
Although each of these features is implemented and tested in the SymEngine
library itself, the \texttt{BatchOptimizer} class provides both a uniform
interface to their classes and a convenient interface for scalar expressions,
as well as tensorial expressions formed using the \texttt{deal.II} tensor and 
symmetric tensor classes.
It, like the \texttt{Expression} class, is also serializable.

The way the batch optimizer may be employed within a user's code is shown
in the pseudo-code below.
As per usual, one would first define some independent variables, and 
subsequently compute some symbolic expressions that are dependent on these 
independent variables. 
These expression could be, for example, scalar  expressions or tensors of 
expressions.
Instead of evaluating these expressions directly, the user would now create an
optimizer to evaluate the dependent functions.
In this example, the selected arithmetic type numerical result will be of type
\texttt{double}, and the \texttt{LLVM} JIT optimizer will be invoked.
It will employ common subexpression elimination and aggressive optimizations
during compilation.
The user then informs the optimizer of all of the independent variables and the
dependent expressions, and invokes the optimization process.
This is an expensive call, as it determines an equivalent code path to evaluate
all of the dependent functions at once.
However, in many cases each evaluation has significantly less computational cost
than evaluating the symbolic expressions directly.
Evaluation is performed when the user constructs a substitution map, giving each
independent variable a numerical representation, and passes those to the 
optimizer.
After this step, the numerical equivalent of the individual dependent expressions
may finally be retreived from the optimizer.

@@

  \item \texttt{step-50} is a program that demonstrates the parallel geometric multigrid features
  in \dealii{} as described in \cite{ClevengerHeisterKanschatKronbichler2019}.
  The problem considered is a variable viscosity Laplace equation and it is solved
  with three different approaches:
  (i) With a matrix-based geometric multigrid based on Trilinos or
  PETSc; (ii) with a matrix-free
  geometric multigrid; (iii) using algebraic multigrid (Trilinos ML).
